More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-12-21 2016-Weak-FH-Bisim/WeakBisim.tex
changed notations in proof of thm 4 (s,t, tgt)

       The proof of the other theorems rely on two main lemmas,
dealing respectively with the decomposition of a composed behaviour
between the context and the internal pNet, and with their recomposition. 
We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:

@@

         \newpage
 \section{Proof of Theorem ~\ref{thm-congr-eq}}
 The proof of Theorem~\ref{thm-congr-eq} exhibits classically a bisimulation relation for 
 a 
 composed system.  It considers then an open transition of $P[Q]_{j_0}$ that should be 
 simulated. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
 of $P[Q]_{j_0}$ and obtain an open transition of $P$ and $Q$; the FH-bisimulation 
 property can 
 be applied  to $Q$ to obtain an equivalent family of open transitions of $Q'$; this 
 family is 
 then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
 $P[Q']_{j_0}$ 
 that will simulate the original one.

 \TODO{Ludo: switch the proof to new notations : s,t instead of 21 s2, and Pred s,t}
 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation $\mathcal{R}'=\{(s_1,s_2|\Pred)|s_1=s'_1\uplus s \land 
 s_2=s'_2\uplus s \land s\in \mathcal{S}_P \land (s_1',s_2'|\Pred)\in\mathcal{R}\}$ 
 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{1 i}^{i \in L_P\uplus L}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
 	{s}_{1 i}^{i \in L_P}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{1 i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{1 i}^{i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post'}
 	{\ostate{s_{1 i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{1 i}'^{\, i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post''}
 	{\ostate{s_{1 i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{1 i}'^{\, i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,'}
 \land {\Pred\,}'')$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{1 i}^{i \in 
 		L}},\ostate{{s}_{2 i}^{i \in L'}}|\Pred)\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{s_{2 i}\xrightarrow{a_{i x}}_i s_{2 i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{s_{2 i}^{i\in L'}} \OTarrow {v_x} \ostate{s_{2 i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{1 i}^{i\in L}},\ostate{s_{2 i x}^{i\in 
 		L'}}|\Pred_{tgt_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{tgt_x}\subst{\Post''}\subst{\Post_{OT_x}}\right)$

 We can now apply Lemma~\ref{lem-compose} on each of the $OT'_x$ together with 
 the transition of $P$ and obtain a new family $OT_x$ of open transitions (where for 
 $i\in L_P$, $s_{2 i}=s_{1 i}$ and $s_{2 i x}=s'_{1 i}$, and for $j\in Holes(P)$, 
 $b_{j x}=b_j$):\\[-2ex]
 \[ P[Q']_{j_0}  
 \models
 {\inferrule*[myfraction=\reddottedrule]
 	{\{s_{2 i}\xrightarrow{a_{i x}}_i {s}_{2 i x}\}^{i\in I_x \uplus 
 			(I\cap L_P)},
 		\{\xrightarrow{b_{j x}}_j\}^{j\in J}, \\
 		\Pred\,'\subst{b_{j_0 x}\leftarrow v_x}\land\Pred_{OT_x},  
 		\Post\,'\uplus \Post_{OT_x} }
 	{\ostate{s_{2 i }^{i\in L'\uplus L_Q}} \OTarrow {v_x}
 		\ostate{{s}_{2 i x}^{i\in L'\uplus L_Q}}}
 }
 \]

 Observe that we used the fact that $J=(J\setminus\Holes(Q)\cup 
 \{j_0\})\setminus\{j_0\}\cup 
 (J\cap\Holes(Q))$. Now we have to verify the conditions for the 
 FH-bisimulation between $OT$ and $OT_x$.
 $\forall x, (\ostate{{s'}_{1 i}^{i\in L_P\uplus L}},\ostate{s_{2 i 
 		x}^{i\in L_P\uplus L'}}|\Pred_{tgt_x})\in 
 \mathcal{R}'$ (by definition of
 $\mathcal{R}'$) and in four steps we get:

 \noindent                        
 \begin{small} $\Pred \land \Pred_{OT} \implies
 	\Pred\land(\Pred\,'\land \Pred\,'')\subst{b_{j_0}\leftarrow v_Q}\\ % reversed 
 	%substitution 
 	\implies 
 	(\Pred\land\Pred\,'')\land \Pred\,'\subst{b_{j_0}\leftarrow v_Q}\\ % $b_{j_0}$ only 
 	%appears 
 	%in $\Pred\,'$
 	\implies  \hspace{-2ex}
 	{\displaystyle\bigvee_{x\in X}
 	\left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 	\Pred_{OT_x}
 	\land v_Q\!=\!v_x \land  
 	\Pred_{tgt_x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)
 }
 \\~\qquad\qquad	\land \Pred\,'\subst{b_{j_0}\leftarrow v_Q} \\
 	\implies
 	 \hspace{-2ex}
 	{\displaystyle\bigvee_{x\in X}
 		\left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 		\Pred\,'\subst{b_{j_0}\leftarrow v_x}\land \Pred_{OT_x}
 		\land v_Q\!=\!v_x \land  
 		\Pred_{tgt_x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)
 	}
 	$\end{small}

@@

 Explanations: first, we reversed the substitution; then we
 restricted the 
 substitution as $b_{j_0}$ only appears in $\Pred\,'$; then we
 applied the property of FH-bisimulation for $P$ and rearranged the
 terms using the fact that $v_Q=v_x$. The obtained formula reaches the goal except for 
 two points:\\[-4.3ex] 
 \begin{itemize}
 	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J\cap\Holes(Q)$  but  
 	adding prerequisite on more variables 
 	does not 
 	change the validity of the formula (those variables are not used).
 	\item Concerning the last term, we need 
 	$\Pred_{tgt_x}\subst{\Post_{OT}}\subst{\Post\,'\uplus \Post_{OT_x}}$ 
 	, i.e.:
 	$\Pred_{tgt_x}\subst{\Post\,'\uplus 
 		\Post\,''}\subst{\Post\,'\uplus \Post_{OT_x}}$. We 
 	can conclude by observing that	$\Pred_{tgt_x}$ does not use any variable of $P$ 
 	and thus $\subst{Post\,'}$ has no effect.
 \end{itemize}	
 This proves the  condition of the FH-simulation, the other direction is 
 similar.\qed

        \newpage
        \section{Proof of Theorem ~\ref{thm-ctxt-eq}}
The proof of Theorem~\ref{thm-ctxt-eq} exhibits  a bisimulation relation for a 
composed system. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
of $P[Q]_{j_0}$ and obtain an open transition of $P$ on which the FH-bisimulation 
property can 
be applied  to obtain an equivalent family of open transitions of $P'$; this family is 
then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
$P'[Q]_{j_0}$ 
that will simulate the original one.

>>>>

\subsection{Composition Lemmas}       
       The proof of the other theorems rely on two main lemmas,
dealing respectively with the decomposition of a composed behaviour
between the context and the internal pNet, and with their recomposition. 
We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:

@@

 \subsection{Proof of Theorem ~\ref{thm-congr-eq}}
 The proof of Theorem~\ref{thm-congr-eq} exhibits classically a bisimulation relation for 
 a 
 composed system.  It considers then an open transition of $P[Q]_{j_0}$ that should be 
 simulated. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
 of $P[Q]_{j_0}$ and obtain an open transition of $P$ and $Q$; the FH-bisimulation 
 property can 
 be applied  to $Q$ to obtain an equivalent family of open transitions of $Q'$; this 
 family is 
 then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
 $P[Q']_{j_0}$ 
 that will simulate the original one.

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
 t=t'\uplus s'' \land s''\in \mathcal{S}_P \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{i}^{i \in L_P\uplus L}},\ostate{{t}_{i}^{i \in L'}\uplus 
 	{s}_{i}^{i \in L_P}}|\Pred_{s,t})\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{i}^{~i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post'}
 	{\ostate{s_{i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{i}'^{~ i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post''}
 	{\ostate{s_{i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{i}'^{~ i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,'}
 \land {\Pred\,}'')$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{i}^{i \in 
 		L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{t_{i}\xrightarrow{a_{i x}}_i t_{i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{t_{i}^{i\in L'}} \OTarrow {v_x} \ostate{t_{i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
 		L'}}|\Pred_{s x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{s x}\subst{\Post''}\subst{\Post_{OT_x}}\right)$

 We can now apply Lemma~\ref{lem-compose} on each of the $OT'_x$ together with 
 the transition of $P$ and obtain a new family $OT_x$ of open transitions (where for 
 $i\in L_P$, $t_{i}=s_{i}$ and $t_{i x}=s'_{i}$, and for $j\in Holes(P)$, 
 $b_{j x}=b_j$):\\[-2ex]
 \[ P[Q']_{j_0}  
 \models
 {\inferrule*[myfraction=\reddottedrule]
 	{\{t_{i}\xrightarrow{a_{i x}}_i {t}_{i x}\}^{i\in I_x \uplus 
 			(I\cap L_P)},
 		\{\xrightarrow{b_{j x}}_j\}^{j\in J}, \\
 		\Pred\,'\subst{b_{j_0 x}\leftarrow v_x}\land\Pred_{OT_x},  
 		\Post\,'\uplus \Post_{OT_x} }
 	{\ostate{t_{i }^{i\in L'\uplus L_Q}} \OTarrow {v_x}
 		\ostate{{t}_{i x}^{i\in L'\uplus L_Q}}}
 }
 \]

 Observe that we used the fact that $J=(J\setminus\Holes(Q)\cup 
 \{j_0\})\setminus\{j_0\}\cup 
 (J\cap\Holes(Q))$. Now we have to verify the conditions for the 
 FH-bisimulation between $OT$ and $OT_x$.
 $\forall x, (\ostate{{s'}_{i}^{~i\in L_P\uplus L}},\ostate{t_{i 
 		x}^{i\in L_P\uplus L'}}|\Pred_{s x})\in 
 \mathcal{R}'$ (by definition of
 $\mathcal{R}'$) and in four steps we get:

 \noindent                        
 \begin{small} $\Pred_{s,t} \land \Pred_{OT} \implies
 	\Pred_{s,t}\land(\Pred\,'\land \Pred\,'')\subst{b_{j_0}\leftarrow v_Q}\\ % reversed 
 	%substitution 
 	\implies 
 	(\Pred_{s,t}\land\Pred\,'')\land \Pred\,'\subst{b_{j_0}\leftarrow v_Q}\\ % $b_{j_0}$ 
 	%only 
 	%appears 
 	%in $\Pred\,'$
 	\implies  \hspace{-2ex}
 	{\displaystyle\bigvee_{x\in X}
 	\left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 	\Pred_{OT_x}
 	\land v_Q\!=\!v_x \land  
 	\Pred_{s x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)
 }
 \\~\qquad\qquad	\land \Pred\,'\subst{b_{j_0}\leftarrow v_Q} \\
 	\implies
 	 \hspace{-2ex}
 	{\displaystyle\bigvee_{x\in X}
 		\left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 		\Pred\,'\subst{b_{j_0}\leftarrow v_x}\land \Pred_{OT_x}
 		\land v_Q\!=\!v_x \land  
 		\Pred_{s x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)
 	}
 	$\end{small}

@@

 Explanations: first, we reversed the substitution; then we
 restricted the 
 substitution as $b_{j_0}$ only appears in $\Pred\,'$; then we
 applied the property of FH-bisimulation for $P$ and rearranged the
 terms using the fact that $v_Q=v_x$. The obtained formula reaches the goal except for 
 two points:\\[-4.3ex] 
 \begin{itemize}
 	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J\cap\Holes(Q)$  but  
 	adding prerequisite on more variables 
 	does not 
 	change the validity of the formula (those variables are not used).
 	\item Concerning the last term, we need 
 	$\Pred_{s x}\subst{\Post_{OT}}\subst{\Post\,'\uplus \Post_{OT_x}}$ 
 	, i.e.:
 	$\Pred_{s x}\subst{\Post\,'\uplus 
 		\Post\,''}\subst{\Post\,'\uplus \Post_{OT_x}}$. We 
 	can conclude by observing that	$\Pred_{s x}$ does not use any variable of $P$ 
 	and thus $\subst{Post\,'}$ has no effect.
 \end{itemize}	
 This proves the  condition of the FH-simulation, the other direction is 
 similar.\qed

        \subsection{Proof of Theorem ~\ref{thm-ctxt-eq}}
 \TODO{Ludo: switch the proof to new notations : s,t instead of s1 s2, and Pred s,t}
The proof of Theorem~\ref{thm-ctxt-eq} exhibits  a bisimulation relation for a 
composed system. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
of $P[Q]_{j_0}$ and obtain an open transition of $P$ on which the FH-bisimulation 
property can 
be applied  to obtain an equivalent family of open transitions of $P'$; this family is 
then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
$P'[Q]_{j_0}$ 
that will simulate the original one.
